From Nested Virtualization Architecture \\To 2D Security Management for Distributed Clouds

### 2014-03-29 paper/draft.tex
commit

\documentclass{sig-alternate}
\usepackage[utf8]{inputenc} 
%\usepackage[hidelinks]{hyperref}
%\usepackage[autostyle,french=guillemets,babel]{csquotes}
\usepackage{graphicx} % support the \includegraphics command and options
%\usepackage[backend=biber]{biblatex}
%\addbibresource{biblio.bib}
% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{booktabs}
\usepackage{array}
\usepackage{paralist}
\usepackage{verbatim} 
\usepackage{amsfonts}
\usepackage{fancyhdr}
\usepackage{times}
%\pagestyle{fancy} % options: empty , plain , fancy
%\renewcommand{\headrulewidth}{0pt} % customise the layout...
%\lhead{}\chead{}\rhead{}
%\lfoot{}\cfoot{\thepage}\rfoot{}

@@

%The first challenge means that threats will be both vertical, spanning layers, and horizontal, spanning domains  the SUPERCLOUD being a multi-layer, multi-domain distributed cloud infrastructure. Vertically, attacks currently target several infrastructure layers (e.g., VMs, hypervisor). Unfortunately, most existing defenses are generally for single layers only. This makes it difficult to grasp the overall extent of an attack. Horizontally, threats may propagate through the network from cloud to cloud. Unfortunately, most existing defenses are for single clouds only. Lack of interoperability in security policies and mechanisms is thus a major barrier to  unified hybrid cloud security. The second challenge means that although many defense mechanisms may be leveraged in this 2D landscape, they remain difficult to configure and to reconcile. The current “by hand” approach becomes a nightmare for security administrators. Despite some automated solutions for managing security of single clouds, there is currently no solution truly applicable to federations of clouds in practice.

@@

%% Motivation
%Cloud computing platform are a solid reality nowadays. 
\noindent Distributed cloud computing is now becoming a reality. Modern cloud platforms are able to provide elastic use of resources, optimizing their allocation to meet user needs. However, currently deployed distributed cloud infrastructures present several limitations.

% Interop
DCC does not handle heterogeneous infrastructures which secludes providers in non-evolving hardware.
% Securite
Also, security remains the last hurdle toward wide adoption of DCC. The user data are no more self contained but distributed among all parties without any proof of confidentiality, integrity or availability.

%% Specific problem considered
We classified problems related to DCC horizontally and vertically. The horizontal view gathers features related to interactions between IaaS stacks, such as live migrations. The vertical view addresses features close to the isolation of execution environment on the same physical machine. This classification provide a global coverage of problems to solve to deliver a secure, stable and robust architecture.

%% "In this paper we show that"
As a first element of answer, the nested virtualization~\cite{turtle:ibm} performed a breakthrough with enhanced isolation capabilities on the same physical machine. The hypervisor is now virtualized by a thin layer~\cite{cloudvisor:zhang} to counter failures. Furthermore, the horizontal challenges were also included in this tiny layer to deliver an homogeneous interface for commodity hypervisors~\cite{art:blan, xclo:blank}. However these classic approaches suffers from several shortcomings detailed in Section~\ref{ref:challenges}. Thus we propose a novel architectural approach to undertake remaining problems.

%% Differences with others
The previous works created both horizontal and vertical basic blocks from scratch. While being a necessary step, these approaches does not inherits the advances performed in terms of operating system security. The software architecture are mainly monolithic and yet-another-layer fends off security problems without
figuring them out. Our approach learns from security issues fixed by OS architecure~\cite{} and apply these patterns to leverage nested virtualization. In this paper we show that component-based architecures are a promising design for the next generation of clouds.

This paper is organized as follows. Section~\ref{ref:challenges} details the constraints of nested virtualization for the two dimension of our solution. In Section~\ref{sec:rw} we analyze the nested virtualization ecosystem and position existing solutions regarding our design principles. In Section~\ref{sec:classic}, we details limitations of classical nested virtualization approach and in Section~\ref{sec:mh} we present the microvisor approach. In Section~\ref{sec:archcomp} we compare cloud architectures propositions, combining nested virtualization and microvisors, leveraging the design principles grid introduced before.
%Furthermore, in Section 6 we define a security architecture for multi-cloud scenario, with an analysis of necessary software components to satisfy vertical design principles.
In Section~\ref{sec:conclu}, we conclude the paper with future works and our vision of the SuperCloud.
% in section in section 6 we define a security architecture for multi-cloud scenario, with an analysis of necessary software components to satisfy vertical design principles. In section 8, we conclude the paper with future works in section 8.

@@

First, deploying cloud applications able to transparently leverage resources of several heterogeneous platforms is still difficult. The concept of super-cloud~\cite{art:blan, xclo:blank} tries notably to overcome such issues. Multi-platform coordination is hard, primarily due to lock-in to different standards such as VM formats. Unfortunately, such lock-in is not only due to market strategic decisions, but also reflects heterogeneity of virtualization technologies. For example, each cloud provider may use different storage systems or different formats to store VMs images. Different hypervisors have nowadays a low degree of compatibility between them. This generally prevents leveraging some popular features like VM live migration between different platforms. Incompatibility is even greater if we consider that each platform may adopt several types of virtualization techniques. This situation induces a tight dependency between the customer and the cloud provider. It may prevent guaranteeing service availability by relying on alternative providers, or to set up geographical optimizations due to cost or origin of traffic. This has resulted in some dramatic data server failures in recent years, causing prolonged outage of service for many applications. In what follows, we refer to this class of multi-platform issues as \textit{horizontal} challenges.

Second, a number of issues have been raised about user data privacy and platform integrity. From the user perspective, users would like to preserve privacy of their data stored on the cloud platform. However, due to overprivileged hypervisors, a curious or malicious administrator could easily retrieve such data~\cite{cloudvisor:zhang}. In a multi-tenant cloud platform, physical resource sharing between users may lead not only to data leakage or but also attacks against the IaaS~\cite{bot:att, sec:you}.
From the provider perspective, a key objective is to thus guarantee platform integrity against rogue VMs. However, commodity general-purpose hypervisors (GPHs) make it difficult to observe deeply all layers of a running IaaS platform. Indeed, running GPHs at the highest level of privilege hampers detection and reaction on the hypervisor itself in presence of faulty or malicious behavior -- unless to cause interruption of service for all hosted VMs. Thus, both user-centric and provider-centric visions of the cloud have a problem with overprivileged hypervisors.
%: the platform administrator cannot perform in-depth infrastructure monitoring nor can users concretely inspect administrator actions on their data. 
A major challenge for new generations of cloud infrastructures is thus to find acceptable trade-offs between such conflicting requirements. In what follows, we refer to this class of layer-oriented issues as \textit{vertical} challenges.

We identified five key design principles (DP) that a IaaS architecture should implement to meet the previous challenges. In what follows, those principles will serve as reading template to analyze different architectures.
\begin{itemize}[]
\item \textbf{DP1 User Data Privacy} A curious or malicious administrator (or another platform user) should not be authorized to inspect and analyze user data and VM instances without explicit user consent relying on mutually trusted services. A trade-off should be found to allow providers to detect and react to potential threats, but also to prevent unlimited platform and user data access in such inappropriate operator behaviors. Thus, a platform administrator should not manage the infrastructure directly using the highest level of privilege. Instead, his actions on user data objects should be validated by a mutually-trusted kernel run in the most privileged mode. This kernel should also enforce memory page encryption to prevent snooping when pages are swapped out.
\item \textbf{DP2 Fail-Safe Modular Architecture} The architecture should be fail-safe: for instance, to enable automatically migration of instances or user-data in presence of a threat and to recover from software failures. To reach this goal, the amount of safety-critical components should be as low as possible to mimize the attack surface. Similarly, the overall design should be as modular as possible to enable platform reconfiguration.
\item \textbf{DP3 Small TCB} The number of potential failures and bugs in the platform is directly linked with the code size executed at the highest level of privilege. Thus, an architecture with a tiny TCB (Trusted Computing Base) will improve safety and integrity by design.
\item \textbf{DP4 Inter-Platform Support} The cloud architecture should allow supporting directly different platforms and service providers. The same user virtual instances should be allowed to be run (and potentially migrated) unmodified over different IaaS platforms.
\item \textbf{DP5 Compatibility with Legacy} Evolution in the IaaS architecture should avoiding disrupting existing third party applications. Thus, platform interfaces should be kept as close as possible to existing ones, to prevent code rewriting notably for applications.
\end{itemize}
In addition to the above functional design principles that specify which features have to be achieved by the platform architecture, another set of non-design principles (NDP) should be considered regarding effective feasibility of implementing the architecture. We just give an example in terms of platform usability:
\begin{itemize}[]
\item \textbf{NDP1 Minimal Additional Code} A major issue could be represented by the amount of code to be added and maintained to an existing architecture to meet functional requirements. Despite not being a strictly a technical limitation of the architecture, it could become a show-stopper to its practical use.
\end{itemize}
In our analysis, we will consider not only the overall suitability of the architecture, but also the amount of effort needed to develop and deploy it. Table \ref{int:des} summarizes the previous principles and which class of requirements they address.

\begin{table}
\caption{Design Principles for a Distributed IaaS Platform.\label{int:des}}
\begin{tabular}{lll}
\toprule
\textbf{Id} & \textbf{Design Principle} & \textbf{Class of Requirement} \\
\midrule
DP1 & User Data Privacy & Vertical \\
DP2 & Fail-Safe Modular Architecture & Vertical \\ 
DP3 & Small TCB & Vertical \\
DP4 & Inter-Platform Support & Horizontal \\
DP5 & Compatibility with Legacy & Horizontal \\
NDP1 & Minimal Additional Code & -- \\
\bottomrule
\end{tabular}
\end{table}

@@

% Intro + performance issues
Several works presented in literature integrates these a subset of this design principles in their architecture, leveraging nested virtualization \cite{turtle:ibm,art:blan, cloudvisor:zhang}
Cloud computing feature set could be enlarged exploiting two layers of virtualization instead of just one. The presence of two different hypervisors allows the possibility to specialize hypervisors functions. Instead of having a general purpose hypervisor we could split the hypervisor role in two different layers, interconnected with a well defined interface.
Each layer will have to care about a subset of features, simplifying overall development.
However, nested virtualization on x86 architecture faced impeding performance issues \cite{rec:virt}.

% Perfomance issuess adressed
However, several works demonstrate that it is possible to put in place low overhead nested virtualization using special approaches \cite{turtle:ibm} . Moreover, other works try to recover performance loss exploiting multi-level consolidation, introduced in nested virtualization \cite{art:blan}. In conclusion, nested virtualization slowdown could be considered nowadays an acceptable price to achieve advanced features and, in addition, hardware manifacturers are developing and proposing new hardware facilities to reduce this overhead \cite{vmcs:nakajima}.

To sum up, no architecture had been presented to address vertical and horizontal class of problems at the same time until now. XenBlanket does not investigate cloud architectural security concerns (DP3,DP2), limiting the focus on new potential feature of a coordinated multi-provider cloud. Cloudvisor proposed a security architecture that leverages nested virtualization without allowing the possibility of executing more than one nested hypervisor at time, that is a crucial requirements of a multi-provider cloud platform (DP4). In a SuperCloud scenario, each user has to run its own L1 hypervisor instance to enforce a compatibility layer, overcoming incompatibilities between different providers.

Other approaches leveraging only one layer of virtualization does not achieve the whole design-principle satisfaction. Self-service cloud approach (SSC) \cite{ssc:art} design a complete security architecture to overcome major security and privacy issues of currently deployed cloud infrastructures, in order to alleviate the different needing user-centric and provider-centric vision of cloud computing presented in \ref{sec:intro}. Even if this architecture reproduces an approach similar to the SuperCloud without leveraging an extra layer of virtualization (DP4), the approach changes the behavior where the user acts introducing new components that have to be directly managed by the user. Therefore, the main drawback of this solution is to change the interface with the user, preventing the possibility to straightforwardly move to the new platform without having to change its applications (DP5).

% Integration NDSS Taming
Furthermore, DeHype \cite{wu2013TamHosHypMosDepExe} presented a dissected version of KVM hypervisor, trying to reduce the TCB of this popular hypervisor. The approach consisted in demoting a great part of the hypervisor codebase to a user-level execution. DeHype satisfies all vertical design principles listed before (DS1,DS2,DS3), reinterpreting the microkernel approach defining  the greatest part of code to userspace. DeHype increases a lot the robustness of a cloud platform leveraging KVM, limiting for example the possibility of attack in multi-tenancy platform, but it does not address horizontal design principles about a multi-provider service.

@@

\subsection{Principle}

\noindent \textit{Nested virtualization (NV)} refers to a system architecture with two layers of virtualization~\cite{turtle:ibm}. It consists in ``virtualized'' virtualization: the guest system virtualizes a nested guest. The concept may be generalized to an arbitrary number of nested guest layers, leading to recursive virtualization~\cite{rec:virt,Popek:1974:FRV:361011.361073}. (Full) virtualization is able to execute tasks in a deprivileged context that normally requires a privileged execution. With NV, this approach may be extended to virtualization itself: a guest hypervisor may be run with nested guest VMs without requiring any change to any of them.
%More formally, Popek \& Golberg \cite{Popek:1974:FRV:361011.361073} discussed this subject, introducing the concept of ``recursively virtualizable'' machine and giving an initial definition of recursive virtual machines: If it is possible for a virtual machine to run under itself a copy of the VMM that could also exhibit all the properties of the VMM and this procedure could be repeated until the resources of the systems are consumed, then the original machine is recursively virtualizable.

Two layers should be distinguished when migrating to an NV architecture: (1) the hypervisor running above the hardware (\textit{L0 hypervisor}); and (2) the nested hypervisor (\textit{L1 hypervisor}). Nested guests are generally referred to as \textit{L2 guests}.
Unfortunately, virtualization comes with a price due to the extra layer of indirection introduced: the computational overhead is generally acceptable when dealing with single-layer virtualization, but grows exponentially when adding further levels~\cite{rec:virt}. This huge performance overhead prevented NV adoption during many years. However, Ben Yehuda et al. finally proposed an architecture tending towards more reasonable overheads~\cite{turtle:ibm}.

NV feasibility opened plenty of new possibilities and avenues for research that have been largely discussed in the literature~\cite{turtle:ibm,rec:virt,art:blan}. Such works all tend to diversify hypervisor functionalities, introducing new features but also keeping existing ones. It turns out that each layer addresses clearly different sets of issues.

L0 tackles vertical aspects of the platform. It aims to guarantee the best performance and strongest level of isolation possible. L0 represents the last line of defense of the platform and a privileged point for monitoring of its global status.

L1 is responsible for horizontal feature set. It has to provide the widest possible support for different platforms and the ability to virtualize using different techniques e.g., para-virtualization, hardware-assisted virtualization, dynamic binary translation. {\sffamily To address such requirements, Williams et al.~\cite{art:blan} proposed a patched version of Xen, Xen Blanket, to realize an L1 virtualization layer designed to run over different hypervisors. This notably enables to run and to perform live migration of the same VM over different physical cloud infrastructures.} % A migrer dans le related work?

@@

\noindent Today, NV presents several limitations related to hardware support {\sffamily and software state of implementation}.

>>>>

\documentclass{sig-alternate}
\usepackage[utf8]{inputenc} 
%\usepackage[hidelinks]{hyperref}
%\usepackage[autostyle,french=guillemets,babel]{csquotes}
\usepackage{graphicx} % support the \includegraphics command and options
%\usepackage[backend=biber]{biblatex}
%\addbibresource{biblio.bib}
% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{booktabs}
\usepackage{array}
\usepackage{paralist}
\usepackage{verbatim} 
\usepackage{amsfonts}
\usepackage{fancyhdr}
\usepackage{times}
\usepackage{url}
%\pagestyle{fancy} % options: empty , plain , fancy
%\renewcommand{\headrulewidth}{0pt} % customise the layout...
%\lhead{}\chead{}\rhead{}
%\lfoot{}\cfoot{\thepage}\rfoot{}

\hyphenation{he-terogeneous infra-structure lea-kage rea-lizing accep-table modula-rity tool-kits ho-rizontally me-chanisms}

@@

@@

%% Motivation
%Cloud computing platform are a solid reality nowadays. 
% Modern cloud platforms are able to provide elastic use of resources, optimizing their allocation to meet user needs. 
% la prochaine mutation majeure du cloud computing est celle du nuage hybride, qui fédère dans une même infrastructure nuages privés, publics, ou communautaires. L’intérêt est de s’affranchir d’un fournisseur de nuage dédié, ou de bénéficier de la sécurité d’un nuage privé, tout en profitant de la flexibilité d’un nuage public.
% User-centric...
% Cloud computing today is essentially still provider-centric (PC). The landscape is an increasing number of fiercely competing cloud service providers (CSP), operating multiple heterogeneous clouds. Each CSP offers its own, feature-rich offerings for customer VMs, with a great diversity of services in areas like security (e.g., system and network intrusion monitoring, firewalling, anti-malware, data security), resource management (e.g., elastic load balancing), or high availability (e.g., checkpointing/recovery). Such services usually come in the form of virtual appliances, often based on hypervisor support.
% A strong push has thus recently been made towards a user-centric (UC) vision of the cloud  It might be described as a paradigm change in delivering cloud resources (e.g., compute, storage). To lift the interoperability barrier,  a resource distribution plane, a.k.a. super-clouds (SCs) is introduced, spanning CSPs to decouple resource production (CSPs) from consumption (customers). To tackle the control challenge, this plane is also an extensibility layer, where each new stakeholder operating the distribution layer (a.k.a. SC provider) may provide its own services enabling the customer to deploy self-service, customizable clouds ranging from the SaaS to the full IaaS level. Technically, the idea is to split the cloud infrastructure into provider- and user-controlled domains or modules, several design alternatives being available such extensible hypervisors, flexible hypervisor management interface (dom0), or nested virtualization.
%This vision brings a number of benefits. First, independence from the provider resulting in lower infrastructure operation overheads, more rapid deployment of services, but also increased homogeneity the customer being free a particular hypervisor technology. Second, increased customizability, as the customer can choose which virtualized services (e.g., for security) may be deployed resulting in fully a la carte clouds. Third, new business opportunities, as SCs and their (virtual) apps might become new markets in the manner of ecosystems from the mobile computing world (e.g., iPhone, Android).

\noindent \textit{Distributed cloud computing (DCC) }is now turning into a reality: modern hybrid cloud platforms enable to benefit seamlessly both from private cloud security and from public cloud flexibility. DCC is also moving user-centric: customers demand greater empowerment to deploy self-service clouds really adapted to their needs, and to move away from a provider-dominated landscape of highly heterogeneous closed platforms and its infamous vendor lock-in~\cite{CSA:TT}.

% Mais pb:
% We classified problems related to DCC horizontally and vertically.
% %% Cclassification des problèmes: Specific problem considered
%The horizontal view gathers features related to interactions between IaaS stacks, such as live migrations. The vertical view addresses features close to the isolation of execution environment on the same physical machine. This classification provide a global coverage of problems to solve to deliver a secure, stable and robust architecture.
However, current \textit{distributed cloud infrastructures (DCI) }suffer from two sets of limitations. These may be broadly classified as \textit{horizontal} and \textit{vertical} challenges.
% Problèmes horizontaux: Mais Interop
%DCC does not handle heterogeneous infrastructures which secludes providers in non-evolving hardware.
%2)	Un manque de vision globale et de flexibilité dans la gestion de la sécurité. En cause : l'hétérogénéité forte des politiques et mécanismes entre nuages élémentaires et leur manque d’interopérabilité. Les plates-formes d’interopérabilité actuelles sont de plus trop monolithiques : elles dépendent fortement de la volonté des fournisseurs de nuages de déployer certains services, et ne conviendront donc pas à tous les types de clients.
% 
%Unfortunately, this vision suffers from a lack of homogeneity. Many of those services have not been deployed due to deficiencies in: (1) unified control; and (2) interoperability. Lack of control is visible through vendor lock-in: such services are so tightly coupled with the CSP and on its willingness to deploy them, with little leeway for the customer. Control is also severely limited by monolithic infrastructures (e.g., non-flexible hypervisors hiding hardware capabilities) preventing fine-grained cloud customization by the customer. Lack of interoperability is mainly due to heterogeneity of services, hypervisor-specific and thus not compatible across CSPs. Despite much efforts, standardization might be too slow or too partial to solve such issues, as both customers and providers might want features that are not in the standard.
%•	Lack of flexibility and control for overall security management: this refines the previous interoperability and control challenges for security. For interoperability, the problem comes from heterogeneity of security components and policies between CSPs and SCs. This has a strong security impact by introducing more vulnerabilities due to mismatching APIs. How also to manage interoperability and trust between SCs? For flexibility: how to clearly separate the infrastructure between UC and PC security management elements? how to make that frontier flexible to achieve a continuum between PC and UC security services? and how to define in that continuum security services that are flexibly composable with other types of services (e.g., replication, transactions)?
Horizontally, DCIs lack interoperability, mainly due to heterogeneity of IaaS services, hypervisor-specific and not compatible across providers. Control is also severely limited by monolithic infrastructures preventing fine-grained cloud customization by users. 
% Problèmes verticaux: Securite
%Also, security remains the last hurdle toward wide adoption of DCIs. The user data are no more self contained but distributed among all parties without any proof of confidentiality, integrity or availability.
% 1)	Un manque de sécurité global. Dans chaque nuage élémentaire formant le nuage hybride, chaque couche logicielle (machines virtuelles, hyperviseur) s'avère extrêmement vulnérable aux attaques, rendant difficile une protection intégrée forte.
% •	Security vulnerabilities in infrastructure layers:  For instance, the hypervisor and its over-privileged dom0 is a target of choice for attackers due to its complexity. This makes difficult an integrated protection. How to place / coordinate counter-measures in those layers that usually are not aware of one another, each with their own monitoring mechanisms, while attacks span several layers? How also to continue to guarantee VM protection when one or several of those layers are compromised? 
Vertically, security remains the last hurdle toward wide adoption of DCIs. Due to their complexity, infrastructure layers remain extremely vulnerable to attacks, making it difficult to achieve strong isolation and integrated protection.

\newpage
%% La NV comme élément de réponse:
% As a first element of answer, the nested virtualization~\cite{turtle:ibm} performed a breakthrough with enhanced isolation capabilities on the same physical machine. The hypervisor is now virtualized by a thin layer~\cite{cloudvisor:zhang} to counter failures. Furthermore, the horizontal challenges were also included in this tiny layer to deliver an homogeneous interface for commodity hypervisors~\cite{art:blan, xclo:blank}. 
% Nested virtualization (NV) brings both a homogeneous interface for distributed hypervisor features and enhanced IaaS security, protecting VMs even in case of hypervisor compromise. 
% The hypervisor is now virtualized by a thin layer
Advances in \textit{nested virtualization (NV)} were a major step forward to tackle such challenges~\cite{turtle:ibm}. By virtualizing the hypervisor through a thin software layer, NV brings enhanced isolation capabilities on the same physical machine, protecting VMs even in case of hypervisor compromise~\cite{cloudvisor:zhang}. NV also provides a distributed homogeneous interface for commodity hypervisors~\cite{art:blan,xclo:blank}.
% Mais..
% - Mais limites...
% However these classic approaches suffers from several shortcomings detailed in Section~\ref{ref:challenges}. 
Practical NV adoption is still hindered by code additions to an already large hypervisor TCB (Trusted Computing Base) and by limited hardware support.  There is also currently no NV architecture solving both vertical and horizontal challenges simultaneously.
% - New architectural challenges are required...
New distributed IaaS architectures are therefore required.

%The previous works created both horizontal and vertical basic blocks from scratch. While being a necessary step, these approaches does not inherits the advances performed in terms of operating system security. The software architecture are mainly monolithic and yet-another-layer fends off security problems withoutfiguring them out. Our approach learns from security issues fixed by OS architecure~\cite{} and apply these patterns to leverage nested virtualization. In this paper we show that component-based architecures are a promising design for the next generation of clouds.

In this paper, we propose a new architectural approach for DCIs. We argue that advances in OS design bring significant improvements to traditional NV architectures: instead of relying on general-purpose hypervisors (GPHs) for both virtualization layers, new kernel designs allow to tackle more efficiently different sets of challenges separately in each layer. We make a position statement that \textit{micro-hypervisor (MH) architectures} combined with NV benefits are a promising design for the next generation of DCIs.

% Contributions.
%1)
%2)
%3)
This paper makes the following contributions. From previous challenges, we derive a set of design principles that a DCI architecture should satisfy. We compare possible architectural designs combining NV with MH features to find the best trade-off between user data privacy, modularity, TCB size, and compatibility with multiple platforms. We finally sketch how the resulting architecture may serve as basis for 2D security management of DCIs: vertically, across layers for hardened IaaS security, and horizontally spanning cloud domains for end-to-end security.
 %The first challenge means that threats will be both vertical, spanning layers, and horizontal, spanning domains  the SUPERCLOUD being a multi-layer, multi-domain distributed cloud infrastructure. Vertically, attacks currently target several infrastructure layers (e.g., VMs, hypervisor). Unfortunately, most existing defenses are generally for single layers only. This makes it difficult to grasp the overall extent of an attack. Horizontally, threats may propagate through the network from cloud to cloud. Unfortunately, most existing defenses are for single clouds only. Lack of interoperability in security policies and mechanisms is thus a major barrier to  unified hybrid cloud security. The second challenge means that although many defense mechanisms may be leveraged in this 2D landscape, they remain difficult to configure and to reconcile. The current “by hand” approach becomes a nightmare for security administrators. Despite some automated solutions for managing security of single clouds, there is currently no solution truly applicable to federations of clouds in practice.

This paper is organized as follows. Section~\ref{ref:challenges} presents the DCI challenges and design principles. Sections~\ref{sec:rw} and~\ref{sec:classic} review the NV ecosystem and analyze the limitations of traditional NV design. Section~\ref{sec:mh} presents the MH approach. Section~\ref{sec:archcomp} compares possible DCI architectures combining NV and MH according to the design principles.
Finally, Section~\ref{sec:arch} presents a preliminary DCI system architecture based on the best resulting design with promising properties for 2D management of DCI security.
%Furthermore, in Section 6 we define a security architecture for multi-cloud scenario, with an analysis of necessary software components to satisfy vertical design principles.
%In Section~\ref{sec:conclu}, we conclude the paper with future works and our vision of the SuperCloud.
% in section in section 6 we define a security architecture for multi-cloud scenario, with an analysis of necessary software components to satisfy vertical design principles. In section 8, we conclude the paper with future works in section 8.

@@

\subsection{Challenges}

\noindent Deploying cloud applications able to transparently use resources of several heterogeneous platforms is still difficult. 
%The concept of super-cloud~\cite{art:blan, xclo:blank} tries notably to overcome such issues. 
Multi-platform coordination is hard, primarily due to lock-in to different standards. Unfortunately, such lock-in is not only due to market strategic decisions, but also reflects heterogeneity of virtualization technologies. For example, each cloud provider may use different storage systems or formats for VM images. Different hypervisors have today a low degree of compatibility between them. This generally hinders using some popular features like VM live migration across different platforms. Incompatibility is even greater considering each platform may adopt several types of virtualization techniques. This situation induces a tight dependency between the customer and the cloud provider. It may prevent guaranteeing service availability by relying on alternative providers, or setting up geographical optimizations according to cost or origin of traffic. This resulted in some dramatic data server failures in recent years, causing prolonged outage of service for many applications. We refer to this class of multi-platform issues as \textit{horizontal challenges}.

Several concerns were also raised about user data privacy and platform integrity. For users, preserving the privacy of their data stored on the cloud platform is important. However, due to overprivileged hypervisors, a curious or malicious administrator may easily retrieve such data~\cite{cloudvisor:zhang}. 
In a multi-tenant cloud, physical resource sharing between users may lead not only to VM data leakage but also to attacks against the IaaS~\cite{bot:att, sec:you}.
For the provider, a primary objective is thus to guarantee platform integrity. Unfortunately, commodity GPHs are not well adapted to observe deeply all layers of a running IaaS platform: running a GPH at the highest privilege level hampers detection and reaction on the hypervisor itself in presence of faulty or malicious behavior. Thus, both users and providers have a problem with overprivileged hypervisors.
%: the platform administrator cannot perform in-depth infrastructure monitoring nor can users concretely inspect administrator actions on their data. 
%A major challenge for new generations of DCIs will be to find acceptable trade-offs between such conflicting requirements. 
We refer to this class of layer-oriented issues as \textit{vertical challenges}.

\subsection{Design Principles}

\noindent We identified the following key DCI design principles to meet the previous challenges: 
%In what follows, those principles will serve as reading template to analyze different architectures.
\begin{itemize}[]
\item \textbf{DP1 User Data Privacy} An administrator (or another platform user) should not be authorized to inspect and analyze user data and VM instances without explicit user consent. A trade-off should be found to allow providers to detect and react to potential threats, but also to prevent unlimited platform and user data access by curious or malicious operators. Thus, a platform administrator should not manage the infrastructure directly using the highest privilege level. Instead, his actions on user data objects should be validated by a mutually-trusted kernel run in the most privileged mode. This kernel should also enforce memory page encryption to prevent snooping when pages are swapped out.
\item \textbf{DP2 Fail-Safe Modular Architecture} The architecture should be fail-safe, e.g., for automatically migration of VMs or user-data during an attack, or to recover from software failures. The number of safety-critical components should minimal to reach a low attack surface. The overall design should also be highly modular to enable platform reconfiguration.
\item \textbf{DP3 Small TCB} The number of vulnerabilities and failures in the platform is directly linked with the code size run at the highest privilege level. A small TCB architecture will improve safety and integrity by design.
\item \textbf{DP4 Inter-Platform Support} The cloud architecture should support different platforms and providers. The user VMs should run (and potentially be migrated) unmodified over different IaaS platforms.
\item \textbf{DP5 Compatibility with Legacy} Evolution in the IaaS architecture should avoiding disrupting existing third party applications. Platform interfaces should be kept as close as possible to existing ones to prevent code rewriting notably for applications.
\end{itemize}
In addition to the previous functional principles, another set of principles should be considered regarding effective feasibility of implementing the DCI architecture. For instance in terms of platform usability:
\begin{itemize}[]
\item \textbf{DP* Minimal Additional Code} Despite not being a strictly a technical limitation, the amount of code to be added and maintained to an existing architecture to meet functional requirements could become a show-stopper to its practical use.
\end{itemize}
In our analysis, we will consider not only the overall suitability of the architecture, but also the amount of effort needed to develop and deploy it. Table \ref{int:des} summarizes the previous principles and the class of challenges they address.

\begin{table}
\caption{DCI Design Principles.\label{int:des}}
\begin{tabular}{lll}
\toprule
\textbf{Id} & \textbf{Design Principle} & \textbf{Class of Challenges} \\
\midrule
DP1 & User Data Privacy & Vertical \\
DP2 & Fail-Safe Modular Architecture & Vertical \\ 
DP3 & Small TCB & Vertical \\
DP4 & Inter-Platform Support & Horizontal \\
DP5 & Compatibility with Legacy & Horizontal \\
DP* & Minimal Additional Code & -- \\
\bottomrule
\end{tabular}
\end{table}

@@

% Intro + performance issues
\noindent A growing body of research has investigated building architectures and infrastructures integrating a subset of those design principles leveraging NV~\cite{turtle:ibm, hvx:usenix, Inception, art:blan, cloudvisor:zhang}. The distributed cloud feature set may be expanded using two levels of virtualization instead of a single one. This design allows to specialize hypervisor functions per layer, each layer addressing a subset of features, simplifying overall development compared to a GPH. NV development on x86 architecture was initially hindered by severe performance issues~\cite{rec:virt}.
% Perfomance issuess adressed
However, several works have shown feasibility of realizing low-overhead NV using dedicated implementation architectures~\cite{turtle:ibm}, or to recover performance loss exploiting multi-level consolidation~\cite{art:blan}. NV slowdown can now be considered an acceptable price for advanced features -- hardware manufacturers are also developing new hardware primitives to reduce this overhead~\cite{vmcs:nakajima}.

However, so far no architecture has addressed both vertical and horizontal set of challenges simultaneously. XenBlanket~\cite{art:blan} does not investigate cloud architectural security concerns (DP2, DP3), limiting its focus to providing new features on a coordinated multi-provider cloud. Cloudvisor~\cite{cloudvisor:zhang} proposed a NV-based security architecture, but cannot run more than one nested hypervisor at a time, which is a critical requirement for a multi-provider cloud platform (DP4). In the super-cloud vision~\cite{xclo:blank}, each user runs its own L1 hypervisor instance to realize  a blanket layer overcoming incompatibilities between different providers. However, security issues have not been addressed.

Other approaches based on a single layer of virtualization do not fulfill all principles either. Self-Service Clouds (SSC)~\cite{ssc:art} introduces a complete security architecture to overcome major security and privacy issues of currently deployed cloud infrastructures. The aim is to reconcile user-centric and provider-centric requirements of the cloud. Despite being similar to the super-cloud approach (DP4), SSC introduces new components that have to be directly managed by the user. This change of interface with user VMs prevents moving straightforwardly to the new platform without having to change its applications (DP5).

% Integration NDSS Taming
To reduce the TCB size, DeHype~\cite{wu2013TamHosHypMosDepExe} presented a dissected version of KVM, moving a great part of the code to user space. DeHype satisfies all vertical design principles, re-interpreting a micro-kernel design in a KVM setting. It increases robustness and security of a cloud platform leveraging KVM (DP1, DP2, DP3), but does not address horizontal challenges for a multi-provider service.

@@

\subsection{Approach}

\noindent \textit{Nested virtualization (NV)} refers to a system architecture with two layers of virtualization: the guest OS virtualizes a nested guest. The concept may be generalized to an arbitrary number of nested guest layers, leading to recursive virtualization~\cite{rec:virt,Popek:1974:FRV:361011.361073}. (Full) virtualization is able to execute tasks in a deprivileged context that normally requires a privileged execution. With NV, this approach may be extended to virtualization itself: a guest hypervisor may be run with nested guest VMs without requiring any change to any of them.
%More formally, Popek \& Golberg \cite{Popek:1974:FRV:361011.361073} discussed this subject, introducing the concept of ``recursively virtualizable'' machine and giving an initial definition of recursive virtual machines: If it is possible for a virtual machine to run under itself a copy of the VMM that could also exhibit all the properties of the VMM and this procedure could be repeated until the resources of the systems are consumed, then the original machine is recursively virtualizable.

Two layers should be distinguished when migrating to an NV architecture: (1) the hypervisor running above the hardware (\textit{L0 hypervisor}); and (2) the nested hypervisor (\textit{L1 hypervisor}). Nested guests are generally referred to as \textit{L2 guests}.
Unfortunately, virtualization comes with a price due to the extra layer of indirection introduced: the computational overhead is generally acceptable when dealing with single-layer virtualization, but grows exponentially when adding further levels~\cite{rec:virt}. This  performance overhead prevented NV adoption during many years until was finally proposed an architecture tending towards more reasonable overheads~\cite{turtle:ibm}.

NV feasibility opened plenty of new possibilities and avenues for research that have been largely discussed in the literature. Such works all tend to diversify hypervisor functionalities, introducing new features but also keeping existing ones. Each layer clearly addresses different sets of issues:
\begin{itemize}
\item L0 deals with vertical aspects of the platform. It aims to guarantee the best performance and strongest level of isolation. L0 represents the last line of defense of the platform and a privileged point for monitoring of its global status.
\item L1 deals with the horizontal feature set. It should provide the widest possible support for different platforms and the ability to virtualize using different techniques e.g., para-virtualization, hardware-assisted virtualization, dynamic binary translation. 
\end{itemize}
%{\sffamily To address such requirements, Williams et al.~\cite{art:blan} proposed a patched version of Xen, Xen Blanket, to realize an L1 virtualization layer designed to run over different hypervisors. This notably enables to run and to perform live migration of the same VM over different physical cloud infrastructures.} % A migrer dans le related work?

@@

\noindent Today, NV presents several limitations related to hardware support and TCB size.
